Outline flowCore paper

### 2008-04-25 paper/outline.tex
paper: first paragraph for the outlook-application section
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@31325 bc3139a8-67e5-0310-9ffc-ced21a209358

\section{Standard flow operations}
The basic operations in flow cytometry analysis are almost always the
same: the data need to be compensated (if not already done on the
instrument side), often transformed and sup-population of interest
need to be selected based on a set of (predominantly sequential)
gates. All software solutions for flow data analysis offer more or
less elaborate support for these kinds of operation, most of them in
an interactive, GUI-driven spirit. The approach we have taken in
flowCore is to abstractly describe these operations and build a set of
tools to perform them on both flowFrames as well as on entire
flowSets. While transformation and to a certain extent also
compensation are fairly routine operations with only limited potential
for improvement, being able to implement new methodology for gating of
flow cytometry data and extending the capabilities of flowCore through
the usual ways of object oriented programming is a feature, that
clearly sets our framework apart from all the other flow toolsets
around. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and access. Third
party methods implemented by others can act on their own as
first-class citizens in the analysis framework, without breaking the
workflow or the basic infrastructure. This design allows for the
straight forward extension of flowCore's capabilities and already
fostered the development of a number of valuable
improvements. \citep{(flowClust, plateCore)}

@@

{\bf NLM:} example: cite flowClust and floViz paper?

@@

{\bf NLM:} We could illustrate/summarize:
\begin{itemize}
\item the power of subsetting: for example to manipulate 
patient below a certain age or of the same sex...
\item Gating: filterSet strategy, kmeanFilter
\item visualization adding metadata as parameter
example: cite flowViz paper
\end{itemize}

>>>>

\section{Standard flow operations}
The basic operations in flow cytometry analysis are almost always the
same: the data need to be compensated (if not already done on the
instrument side), often transformed and sup-population of interest
need to be selected based on a set of (predominantly sequential)
gates. All software solutions for flow data analysis offer more or
less elaborate support for these kinds of operation, most of them in
an interactive, GUI-driven spirit. The approach we have taken in
flowCore is to abstractly describe these operations and build a set of
tools to perform them on both flowFrames as well as on entire
flowSets. While transformation and to a certain extent also
compensation are fairly routine operations with only limited potential
for improvement, being able to implement new methodology for gating of
flow cytometry data and extending the capabilities of flowCore through
the usual ways of object oriented programming is a feature, that
clearly sets our framework apart from all the other flow toolsets
around. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and access. Third
party methods implemented by others can act on their own as
first-class citizens in the analysis framework, without breaking the
workflow or the basic infrastructure. This design allows for the
straight forward extension of flowCore's capabilities and already
fostered the development of a number of valuable
improvements. \citep{lo2008agf} (+plateCore).

@@

@@

{\bf NLM:} Do we really need supplementary material, can you cite the
paper and packages's vignette?

The flowCore package has already been used to analyze a number of
datasets \citep{gasparetto2004ice,brinkman2007hcf}. Its concepts have
also been integrated in other Bioconductor packages. As an example,
\cite{lo2008agf} have recently proposed an automatic gating approach
via robust model-based clustering using the flowCore data model as the
container for the experimental data and some of the concepts as
clustering methods. The authors have wrapped their clustering method
in a filter operation and the outcome extends the flowCore
filterResult class. Most of the methods available for the flowCore
filterResult class are then available to manipulate the clustering
result. Similarly, the flowViz package make use of the filter concepts
to group variables in a display \cite{sarkar2008ufv}. Even more, the
flowViz plotting functions take advantages of the flowCore data model
to represent multiple variables within one display. To that aim and to
list a few, the plotting functions make use of the data structure to
modify the scale, group, subset, and color the variables.

{\bf NLM:} Should we add some figure or code?
